---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo=TRUE}
if (!file.exists("activity.csv")) {
  download <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  tempfile <- "temp.zip"
  download.file(download, tempfile, method = "curl")
  unzip(tempfile)
}  

activity <- read.csv("activity.csv")
``` 


## What is mean total number of steps taken per day?
```{r, echo = TRUE}
applied <- aggregate(steps ~ date, data = activity, FUN="sum", na.rm=TRUE)
hist(applied$steps)
```

### Calculate Mean and Median
```{r, echo = TRUE}
datamedian <- median(applied$steps)
datamean <- mean(applied$steps)
```

The mean is `r datamean` and median is `r datamedian`

## What is the average daily activity pattern?
```{r, echo = TRUE}
meanapplied <- aggregate(steps ~ interval, data = activity, FUN="mean", na.rm=TRUE)
plot(meanapplied$interval, meanapplied$steps, type="l")

max_interval <- meanapplied[which.max(meanapplied$steps),1]
```

The interval with the highest average steps is `r max_interval`

## Imputing missing values

```{r, echo = TRUE}
narows <- nrow(activity[is.na(activity$steps), ])



```

There are `narows` NA rows in the activity dataset.

## Are there differences in activity patterns between weekdays and weekends?
